---
title: "Homework #1"
format: html
editor: visual
author: "Yefrid Cordoba"
---

```{r}

```

## Problem 1

A physician wanted to estimate the mean length of time µ that a patient had to wait to see him after arriving at the office. A random sample of 36 patients showed a mean waiting time of 23.4 minutes and a standard deviation of 7.2 minutes. Find a 96% confidence interval for µ.\
**R/**

```{r}
n1 <- 36
x_bar1 <- 23.4
sd1 <- 7.2
t_star <- qnorm(0.98) #for n>30 and unknown sigma, we use standard normal distribution
SE <- sd1 / sqrt(n1)
Upper <- x_bar1 + t_star * SE
Lower <- x_bar1 - t_star * SE
Upper
Lower
```

We are 96% confident that the true mean waiting time at the physician's office is between (`r round(Upper, 2)`, `r round(Lower, 2)`) minutes.

## Problem 2

It is hoped that a newly developed pain reliever will more quickly produce perceptible reduction in pain to patients after minor surgeries than a standard pain reliever. The standard pain reliever is known to bring relief in an average of 3.5 minutes. To test whether the new pain reliever works more quickly than the standard one, 50 patients with minor surgeries were given the new pain reliever and their times to relief were recorded. The experiment yielded sample mean x ̅=3.1 minutes and sample standard deviation s = 1.5 minutes. Is there sufficient evidence in the sample to indicate, at the 5% level of significance, that the newly developed pain reliever does deliver perceptible relief more quickly?

```{r}
mu_0 <- 3.5 #standard pain reliever time
n2 <- 50
x_bar2 <- 3.1
sd2 <- 1.5
```

$$H_o : \mu_2 = \mu_0$$
$$H_a  : \mu_2 < \mu_0$$
```{r}
t_obs <- (x_bar2 - mu_0) / (sd2 / sqrt(n2))
t_obs
t_s <- qt(0.05, n2 - 1)
t_s
t_obs < t_s 
pt(t_obs, n2-1, lower.tail = T) < 0.05 #alpha
```
As $t_{obs} < t_s$ is true for a lower tailed test, we reject the null hypothesis, and we have strong evidence to conclude that the recovery time for the new pain reliever is shorter than the standard pain reliever.

## Problem 3
A software company markets a new computer game with two experimental packaging designs. Design 1 is sent to 11 stores; their average sales the first month is 52 units with sample standard deviation 12 units. Design 2 is sent to 6 stores; their average sales the first month is 46 units with sample standard deviation 10 units. 

a. Construct a point estimate and a 95% confidence interval for the difference in average monthly sales between the two package designs assuming σ_1^2=σ_2^2.
         

b. Test at the 1% level of significance (assuming σ_1^2=σ_2^2 ) whether the data provide sufficient evidence to conclude that the mean sales per month of the two designs are different. Use the critical value approach.
